---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rsprefs

<!-- badges: start -->
<!-- badges: end -->

The goal of rsprefs is to help you manage your RStudio preferences.

## Installation

You can install the development version of rsprefs from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("gadenbuie/rsprefs")
```

## Quick Start

1. Snapshot your RStudio preferences to a new GitHub gist.
   Set `path = "new private gist"` to keep your preferences private.
   
   ```r
   rsprefs::rs_prefs_snapshot(name = "sync", path = "new gist")
   ```
   ```
   ✓ Created gist with id 96e8c7f0a624bfed8018987185feb30b
   ℹ Use this gist automatically by adding the following to your ~/.Rprofile.
   Call `usethis::edit_r_profile()`     
  
   options(rsprefs.gist_id = "96e8c7f0a624bfed8018987185feb30b")
  
   ✓ Creating gist with user-prefs.json ... done
   ✓ Writing snapshot sync to new gist ... done
   ```
  
2. Open your `~/.Rprofile` using `usethis::edit_r_profile()` and add the `rsprefs.gist_id` option from the output above.

3. If you change your preferences you can update your snapshot by name, using the gist ID found from the global option.

   ```r
   rsprefs::rs_prefs_snapshot("sync", overwrite = TRUE)
   ```
   
4. On a different computer or session, setup the `rsprefs.gist_id` option and then apply your snapshot.

   ```r
   rsprefs::rs_prefs_snapshot_apply("sync")
   ```

## Manage Your RStudio Preferences

rsprefs stores snapshots of your RStudio preferences.
A snapshot can contain all of your preferences, or just a small selection of preferences.
Snapshots are named and can be stored locally or in a [GitHub gist](https://gist.github.com).

### Saving your RStudio preferences locally

By default, rsprefs will store snapshots in a JSON file in `rappdirs::user_config_dir("rsprefs")`,
but you can choose where to store snapshots with the `path` argument.

In these examples, I'll use a temporary file for demonstration purposes.
Here I'll take a snapshot of my current preferences.

```{r}
tmp_snap <- "tmp_snap.json"

rsprefs::rs_prefs_snapshot("demo", path = tmp_snap)
```

You can list available preference snapshots with `rs_prefs_snapshot_list()`

```{r}
rsprefs::rs_prefs_snapshot_list(path = tmp_snap)
```

or you can apply the preferences in a snapshot with `rs_prefs_snapshot_apply()`.
(You can also set `preview = TRUE` to preview the preferences included in the snapshot without applying them.)

```{r local-prefs-preview}
rsprefs::rs_prefs_snapshot_apply(
  name = "demo", 
  path = tmp_snap, 
  preview = TRUE
)
```

### Choose which preferences to include in the snapshot

Snapshots don't need to include _every_ preference in RStudio.
Instead, you can use snapshot to store specific groups of preferences.
There are three arguments that control which preferences are included:

- `include`: Names of RStudio preferences to include.
  If provided, only these preferences are included.
- `exclude`: Names of RStudio preferences to exclude from the snapshot.
- `source`: The source of the current preference value.
  Preferences are set at different levels, from lowest to highest precedence:

  - `"default"` are RStudio's built-in defaults
  - `"computed"` are detected or supplied from external sources, e.g. the
    path to `git` on your system
  - `"system"` are derived from system-wide `rstudio-prefs.json`
  - `"user"` are set by the user for all sessions (global options)
  - `"project"` are set by the current project session
  
  The default is `"user"`, since these are the preferences you've most
  intentionally set for your IDE. Also, `"project"` preferences are already
  stored in the `.Rproj` file.
  
For example, you could store your favorite pane layout as `"favorite_panes"`:

```{r local-prefs-panes}
rsprefs::rs_prefs_snapshot(
  name = "favorite_panes", 
  path = tmp_snap,
  include = "panes"
)

rsprefs::rs_prefs_snapshot_apply(
  name = "favorite_panes",
  path = tmp_snap,
  preview = TRUE
)
```
  
## Explore preferences

You can obtain a list of available preferences for your version of RStudio using:

```{r}
rsprefs::rstudio_prefs_schema(quiet = TRUE)
```

Or you can use the preferences for the latest version

```{r}
rsprefs::rstudio_prefs
```

This list contains detailed information about every preference, so you can use autocomplete to explore the preferences.

```{r}
rsprefs::rstudio_prefs$highlight_r_function_calls
```

```{r}
rsprefs::rstudio_prefs$insert_native_pipe_operator
```

The `rstudio_prefs` object also comes with helper functions to get, set and toggle preferences:

```{r get-set-toggle-helpers}
# $get() current preference value
original <- rsprefs::rstudio_prefs$insert_native_pipe_operator$get()
original

# $toggle() boolean preferences
rsprefs::rstudio_prefs$insert_native_pipe_operator$toggle()

# $set() the preference directly
rsprefs::rstudio_prefs$insert_native_pipe_operator$set(original)

# check that we've come full circle
rsprefs::rstudio_prefs$insert_native_pipe_operator$get()
```

```{r include=FALSE}
fs::file_delete(tmp_snap)
```
