---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rsprefs

<!-- badges: start -->
<!-- badges: end -->

The goal of rsprefs is to help you manage your RStudio preferences.

## Installation

You can install the development version of rsprefs from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("gadenbuie/rsprefs")
```

```{r library}
library(rsprefs)
```

## Quick Start

1. Snapshot your RStudio preferences to a new GitHub gist.
   Set `path = "new private gist"` to keep your preferences private.
   
   ```r
   snapshot_prefs_save(name = "sync", path = "new gist")
   ```
   ```
   ✓ Created gist with id 96e8c7f0a624bfed8018987185feb30b
   ℹ Use this gist automatically by adding the following to your ~/.Rprofile.
   Call `usethis::edit_r_profile()`     
  
   options(rsprefs.gist_id = "96e8c7f0a624bfed8018987185feb30b")
  
   ✓ Creating gist with user-prefs.json ... done
   ✓ Writing snapshot sync to new gist ... done
   ```
  
2. Open your `~/.Rprofile` using `usethis::edit_r_profile()` and add the `rsprefs.gist_id` option from the output above.

3. If you change your IDE preferences you can update your snapshot by name,
   using the gist ID found from the global option.

   ```r
   snapshot_prefs_save("sync", overwrite = TRUE)
   ```
   
4. On a different computer or session, setup the `rsprefs.gist_id` option and then apply your snapshot.

   ```r
   snapshot_prefs_use("sync")
   ```

## Manage Your RStudio Preferences

rsprefs stores snapshots of your RStudio preferences.
A snapshot can contain all of your preferences, or just a small selection of preferences.
Snapshots are named and can be stored locally or in a [GitHub gist](https://gist.github.com).

### Saving your RStudio preferences locally

By default, rsprefs will store snapshots in a JSON file in `rappdirs::user_config_dir("rsprefs")`,
but you can choose where to store snapshots with the `path` argument.

In these examples, I'll use a temporary file for demonstration purposes.
Here I'll take a snapshot of my current preferences.

```{r}
tmp_snap <- "tmp_snap.json"

snapshot_prefs_save("demo", path = tmp_snap)
```

You can list available preference snapshots with `snapshot_prefs_list()`

```{r}
snapshot_prefs_list(path = tmp_snap)
```

or you can apply the preferences in a snapshot with `snapshot_prefs_use()`.
(You can also set `preview = TRUE` to preview the preferences included in the snapshot without applying them.)

```{r local-prefs-preview}
snapshot_prefs_use(
  name = "demo", 
  path = tmp_snap, 
  preview = TRUE
)
```

If you apply the snapshot and change your mind, you can roll back the changes from the snapshot by calling `snapshot_prefs_undo()`.

### Choose which preferences to include in the snapshot

Snapshots don't need to include _every_ preference in RStudio.
Instead, you can use snapshot to store specific groups of preferences.
There are three arguments that control which preferences are included:

- `include`: Names of RStudio preferences to include.
  If provided, only these preferences are included.
- `exclude`: Names of RStudio preferences to exclude from the snapshot.
- `source`: The source of the current preference value.
  Preferences are set at different levels, from lowest to highest precedence:

  - `"default"` are RStudio's built-in defaults
  - `"computed"` are detected or supplied from external sources, e.g. the
    path to `git` on your system
  - `"system"` are derived from system-wide `rstudio-prefs.json`
  - `"user"` are set by the user for all sessions (global options)
  - `"project"` are set by the current project session
  
  The default is `"user"`, since these are the preferences you've most
  intentionally set for your IDE. Also, `"project"` preferences are already
  stored in the `.Rproj` file.
  
For example, you could store your favorite pane layout as `"favorite_panes"`:

```{r local-prefs-panes}
snapshot_prefs_save(
  name = "favorite_panes", 
  path = tmp_snap,
  include = "panes"
)

snapshot_prefs_use(
  name = "favorite_panes",
  path = tmp_snap,
  preview = TRUE
)
```

### Reset RStudio's Defaults

Occasionally, you might want to reset all of your preferences to RStudio's "factory defaults", for example when teaching or working with a beginner.
To quickly reset all defaults, call

```{r eval=FALSE}
prefs_reset_defaults()
```

If you want to return to your personal preferences, you can undo the change to the defaults with 

```{r eval=FALSE}
snapshot_prefs_undo()
```

For a longer-term return to RStudio's defaults, make sure that you've taken [a snapshot of your personal preferences](#quick-start) before you reset.
  
## Explore preferences

You can obtain a list of available preferences for your version of RStudio using:

```{r}
schema <- prefs_schema()
```

rsprefs includes the preference information for RStudio releases since 1.3.959.
If you're using dailies or preview versions fo RStudio, rsprefs will attempt to download the schema that matches your version, falling back to the latest version if one is not available.

For quick reference, the latest preference details are included in the `rstudio_prefs` object.

```{r rstudio_prefs}
rstudio_prefs
```

This list contains detailed information about every preference, so you can use autocomplete to explore the preferences.

```{r}
rstudio_prefs$highlight_r_function_calls
```

```{r}
rstudio_prefs$insert_native_pipe_operator
```

The `rstudio_prefs` object also comes with helper functions to get, set and toggle preferences:

```{r get-set-toggle-helpers}
# $get() current preference value
original <- rstudio_prefs$insert_native_pipe_operator$get()
original

# $toggle() boolean preferences
rstudio_prefs$insert_native_pipe_operator$toggle()

# $set() the preference directly
rstudio_prefs$insert_native_pipe_operator$set(original)

# check that we've come full circle
rstudio_prefs$insert_native_pipe_operator$get()
```

```{r include=FALSE}
fs::file_delete(tmp_snap)
```
